---
title: "BST 213 HW 3"
author: "Julie Lee"
date: "2022-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, libraries}
# Importing libraries 
library(ggplot2)
library(readr)
library(tidyverse) 
library(readxl) #To read excel files
library(sasLM) # for running like SAS
library(mvtnorm)
library(rstatix) # for tidy table result
```



```{r tidy table}
rm(list = ls())
lbw <- read_excel("lbw.xls")
view(lbw)
summary(lbw)
#tidy data
#race 1 = White, race 2 = Black, race 3 = Other
#history of hypertension 0 = no, 1 = yes
lbw <- lbw |> mutate(ht_des = case_when(ht == 0~0,
                                        ht == 1~1),
                     race_des = case_when(race == 1~0, race == 2~1, race ==3~2))
lbw$ht_des <- as.factor(lbw$ht_des)
lbw$race_des <- as.factor(lbw$race_des)
#make White as reference (race_des) -> biggest sample size
```

## Question 1
Use PROC REG and dummy-coding (see assignment #2) to run a regression with LWT, HT and Race simultaneously as predictors.  Interpret the results and compare to your earlier 1-predictor models.

```{r}
REG(bwt~lwt+ht_des+race_des, Data = lbw) #how SAS runs it
model <- lm(bwt~lwt + ht_des + race_des, data=lbw)
summary(model)
fit <- lm(bwt~lwt + ht_des + race_des, data=lbw)
tidy(fit, conf.int=TRUE) ## show results as tidy table
```


## Question 2
Use PROC GLM and the class statement (see assignment #2) to run a regression with LWT, HT and Race simultaneously as predictors.  Interpret the results and compare to the regression model above

```{r set other to reference race for GLM in SAS}
#changing the reference race to Other Race (bigger sample compare to black)
#White = 1, Black = 2, Other = 0
lbw <- lbw |> mutate(race_des2 = case_when(race == 1~1, race == 2~2, race ==3~0))
lbw$race_des2 <- as.factor(lbw$race_des2)
GLM(bwt~lwt + ht_des + race_des2, Data = lbw, BETA = TRUE, EMEAN = TRUE) # run like SAS
modelgl <- lm(bwt~lwt + ht_des + race_des2, data=lbw)
tidy(modelgl, conf.int=TRUE)
summary(modelgl)
anomodel <- aov(bwt ~ lwt + ht_des + race_des2, data = lbw)
summary(anomodel)
```

