---
title: "BST213_HW2_SK"
author: "Soomin Kim"
date: "2022-09-19"
output: html_document
---

```{r}
library(haven)
library(ggplot2)
```


```{r}
library(readxl)
lbw <- read_excel("~/Google Drive/MPH Harvard Chan/Fall/BST 213/lbw.xls")
View(lbw) 
```


###1.	Use PROC REG to run a model that includes maternal weight as a continuous predictor of birth weight.  Interpret the regression portion (i.e., the bottom) of the printout.   Compare the regression results to a Pearson correlation coefficient. Use PROC GLM to run the same model.  Interpret and compare the results to the regression results from above.

```{r}
lm1 <- lm(bwt~lwt, lbw)
summary(lm1)
confint(lm1)
```
# bwt-hat = 2367.767 + 4.445*lwt-hat
# Mothers who weigh 1 pound more in last menstrual cycle have babies weighing 4.445g more compared to babies of mothers who weigh 1 pound less in last menstrual cycle.

```{r}
ggplot(data=lbw, aes(lwt,bwt))+
  geom_point()+
  geom_smooth(method='lm')
```


```{r}
cor.test(lbw$lwt,lbw$bwt)
```
# pearson r squared = multiple R^2 in lin regression
# p-value for pearson and beta coefficient are the same

###2.	Use PROC REG to run a model that includes history of hypertension as a binary predictor of birth weight. Interpret the regression portion of the printout. Compare the regression results to a 2-sample t-test. Use PROC GLM to run the same model.  Interpret and compare the results to the regression results from above.

```{r}
lbw$ht <- factor(lbw$ht)
lm2 <- lm(bwt~ht, lbw)
summary(lm2)
confint(lm2)
```
# bwt-hat = 2972.42 - 435.67*ht-hat
# Mothers with hypertension have babies weighing 435.67g less compared to babies of mothers without hypertension.

```{r}
ggplot(data=lbw, aes(as.integer(ht),bwt))+
  geom_point()+
  geom_smooth(method='lm')
```


```{r}
t.test(lbw$bwt~lbw$ht, var.equal=TRUE)
```
# p-value for t-test is the same as from lin regression

###3.	Use PROC REG to run a model that includes race (in 3 categories) as a categorical predictor of birth weight.  Interpret the regression portion of the printout. Compare the regression results to a standard ANOVA analysis.  Use PROC GLM to run the same model.  Interpret and compare the results to the regression results from above.

```{r}
lbw$race <- factor(lbw$race)
lm3 <- lm(bwt~race, lbw)
summary(lm3)
confint(lm3)
```
# bwt-hat = 3103.95 - 384.26*Black-hat - 299.93*Other-hat
# Black mothers have babies weighing 384.26g less than babies of White mothers
# Mothers of other races have babies weighing 299.93g less than babies of White mothers

```{r}
ggplot(data=lbw, aes(race,bwt))+
  geom_boxplot()
```


```{r}
aov1 <- aov(bwt~race,lbw)
summary(aov1)
```
# ANOVA p-value same as F stat p-value from lin regression
