---
title: "BST213_HW3_SK"
author: "Soomin Kim"
date: "2022-09-26"
output: html_document
---

```{r}
library(haven)
library(ggplot2)
library(tidyverse)
library(readxl)
lbw <- read_excel("~/Google Drive/MPH Harvard Chan/Fall/BST 213/lbw.xls")
View(lbw)
```

### 1.	Use PROC REG and dummy-coding (see assignment #2) to run a regression with LWT, HT and Race simultaneously as predictors.  Interpret the results and compare to your earlier 1-predictor models.

```{r}
lm1 <- lm(bwt~ lwt + factor(ht) + factor(race), data=lbw)
summary(lm1)
confint(lm1)
```

# Holding hypertension and race constant, one unit increase in lwt of the mother predicts 5.744g higher bwt of the baby
# Holding lwt and race constant, hypertensive mothers are predicted to have a baby with 558.536g less body weight compared to non-hypertensive mothers
# Holding lwt and hypertension constant, black mothers are predicted to have a baby with 433.656g less body weight compared to white mothers
# Holding lwt and hypertension constant, mothers of other races are predicted to have a baby with 226.536g less body weight compared to white mothers
