---
title: "BST 213 HW 4"
author: "Julie Lee"
date: "2022-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries}
# Importing libraries 
library(ggplot2)
library(readr)
library(tidyverse) 
library(readxl) #To read excel files
library(sasLM) # for running like SAS
library(mvtnorm)
library(rstatix) # for tidy table result
library(lsmeans) # least square mean package
```


```{r, tidy tables}
rm(list = ls())
lbw <- read_excel("lbw.xls")
view(lbw)
summary(lbw)
#tidy data
#race 1 = White, race 2 = Black, race 3 = Other
#history of hypertension 0 = no, 1 = yes
#history of smoking 0 = no, 1 = yes
#premature labor 0 = none, 1 = one, 2 = two or more
#presence of uterine irritability 0 = no, 1 yes
#number of physician visits during the first trimester, 0 = none, 1 = one, 2 = two or more
lbw <- lbw |> mutate(ht_des = case_when(ht == 0~0,
                                        ht == 1~1),
                     race_des = case_when(race == 1~0, race == 2~1, race ==3~2))
lbw$ht_des <- as.factor(lbw$ht_des)
lbw$race_des <- as.factor(lbw$race_des)
lbw$smoke <- as.factor(lbw$smoke)
lbw$ptl <- as.factor(lbw$ptl)
lbw$ui <- as.factor(lbw$ui)
lbw$ftv <- as.factor(lbw$ftv)
#make White as reference (race_des) -> biggest sample size
# white race = 0, black race = 1, other race = 2
```


## Question A

Use Proc GLM to get the “Least Squares Means” for HT and Race in a model that includes LWT, HT, and Race.  Interpret the results.

```{r, set other race to reference race for GLM, use race_des2}
#changing the reference race to Other Race (bigger sample compare to black)
#White = 1, Black = 2, Other = 0
lbw <- lbw |> mutate(race_des2 = case_when(race == 1~1, race == 2~2, race ==3~0))
lbw$race_des2 <- as.factor(lbw$race_des2)
GLM(bwt~lwt + ht_des + race_des2, Data = lbw, BETA = TRUE, EMEAN = TRUE) # run like SAS
modelgl <- lm(bwt~lwt + ht_des + race_des2, data=lbw)
tidy(modelgl, conf.int=TRUE)
summary(modelgl)
anomodel <- aov(bwt ~ lwt + ht_des + race_des2, data = lbw)
summary(anomodel)
lsmeans(modelgl, "ht_des")
lsmeans(modelgl, "race_des2")
#library(modelbased) -> can also calcuate least square mean
#estimate_means(modelgl)
```

## Question B

Use Proc Reg to run the model with LWT, HT, and Race (dummy-coded) to get standardized beta’s and partial correlations for all of the predictors.  Interpret the results.
```{r, white as reference race}
# white race = 0, black race = 1, other race = 2
REG(bwt~lwt+ht_des+race_des, Data = lbw) #how SAS runs it
model <- lm(bwt~lwt + ht_des + race_des, data=lbw)
summary(model)
fit <- lm(bwt~lwt + ht_des + race_des, data=lbw)
tidy(fit, conf.int=TRUE) ## show results as tidy table
```

## Question C

Use Proc Reg and start with a model that includes only Smoke.  Then, add to this model each of the other variables (i.e., Smoke and Lwt;  Smoke and Age; etc.) and determine their role (Confounder? Collinear variable?  Predictor?).

```{r, use white race as reference}
REG(bwt~smoke + lwt, Data = lbw) #how SAS runs it
#for smoke
fit0 <- lm(bwt~smoke, data=lbw)
tidy(fit0, conf.int=TRUE)
#for smoke and lwt
fit1 <- lm(bwt~smoke + lwt, data=lbw)
tidy(fit1, conf.int=TRUE) 
#change in Beta after adding maternal weight
(-281.44) - (-268.70)
((-281.44) - (-268.70)) / (-281.44)
#for bwt and lwt
fit1b <- lm(bwt~lwt, data=lbw)
tidy(fit1b, conf.int=TRUE)
#for smoke and maternal weight
fit1c <- lm(as.numeric(smoke)~lwt, data=lbw)
tidy(fit1c, conf.int=TRUE) 
# for smoke and age
fit2 <- lm(bwt~smoke + age, data=lbw)
tidy(fit2, conf.int=TRUE)
#change in beta after adding age
(-281.44) - (-276.03)
((-281.44) - (-276.03)) / (-281.44)
#for bwt and age
fit2b <- lm(bwt~age, data=lbw)
tidy(fit2b, conf.int=TRUE)
# for smoke and age
fit2c <- lm(as.numeric(smoke)~age, data=lbw)
tidy(fit2c, conf.int=TRUE)
# for smoke and race
fit3 <- lm(bwt~smoke + race_des, data=lbw)
tidy(fit3, conf.int=TRUE)
#change in beta after adding race
(-281.44) - (-426.99)
((-281.44) - (-426.99)) / (-281.44)
# for bwt and race
fit3b <- lm(bwt~race_des, data=lbw)
tidy(fit3b, conf.int=TRUE)
# for smoke and premature labor
fit4 <- lm(bwt~smoke + ptl, data=lbw)
tidy(fit4, conf.int=TRUE)
#change in beta after adding premature labor
(-281.44) - (-240.42)
((-281.44) - (-240.42)) / (-281.44)
# for bwt and premature labor
fit4b <- lm(bwt~ptl, data=lbw)
tidy(fit4b, conf.int=TRUE)
#check global test for premature labor and bwt
summary(aov(bwt ~ ptl, data = lbw))
# for smoke and hypertension
fit5 <- lm(bwt~smoke + ht_des, data=lbw)
tidy(fit5, conf.int=TRUE)
#change in beta after adding hypertension
(-281.44) - (-278.57)
((-281.44) - (-278.57)) / (-281.44)
# for bwt and hypertension
fit5b <- lm(bwt~ht_des, data=lbw)
tidy(fit5b, conf.int=TRUE)
# for smoke and uterine irritability
fit6 <- lm(bwt~smoke + ui, data=lbw)
tidy(fit6, conf.int=TRUE)
#change in beta after adding uterine irritability
(-281.44) - (-256.18)
((-281.44) - (-256.18)) / (-281.44)
# for smoke and uterine irritability
fit6b <- lm(bwt~ui, data=lbw)
tidy(fit6b, conf.int=TRUE)
# for smoke and number of physician visits
fit7 <- lm(bwt~smoke + ftv, data=lbw)
tidy(fit7, conf.int=TRUE)
#change in beta after adding number of physician visits
(-281.44) - (-247.50)
((-281.44) - (-247.50)) / (-281.44)
# for bwt and number of physician visits
fit7b <- lm(bwt~ftv, data=lbw)
tidy(fit7b, conf.int=TRUE)
#check global test for number of physician visits and bwt
summary(aov(bwt ~ ftv, data = lbw))
#check global test for number of physician visits and smoke
summary(aov(as.numeric(smoke) ~ ftv, data = lbw))
```


##Question D
Answer the following hypothetical question:  Your current model includes Smoking and Hypertension, both of which are significant.  When you add Age to this model, you find that Age confounds the effect of Smoking (which remains significant), but Age is collinear with Hypertension (which loses significance, although its coefficient is unaffected).  Would you keep Age in the model?

No, because it decreases the power

```{r}
# smoking and hypertension
fit8 <- lm(bwt~smoke + ht_des, data=lbw)
tidy(fit8, conf.int=TRUE)
# age confound smoking
fit9 <- lm(bwt~smoke + ht_des + age, data=lbw)
tidy(fit9, conf.int=TRUE)
# age to the outcome
fit10 <- lm(bwt~age, data=lbw)
tidy(fit10, conf.int=TRUE)
# age to the smoking
fit11 <- lm(as.numeric(smoke)~age, data=lbw)
tidy(fit11, conf.int=TRUE)
#age to hypertension
fit12 <- lm(ht~age, data=lbw)
tidy(fit12, conf.int=TRUE)

```

