---
title: "rlk_bst213_hw5"
author: "Rebecca Kelly"
date: "2022-10-06"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(fig.width = 4, fig.asp = 0.618, echo = FALSE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(tableone)
library(confintr)
library(haven)
library(sasLM)
library(lm.beta)
library(GGally)
```


```{r data import, include=FALSE}

#Import dataset
df<-(read_csv("lbw.csv"))
#Create categorical variables for race, smoking, hypertension, uterine irritability 
# mutate variables to create categorical variables with correct reference group for lm command. 
#collapse ftv and ptl as categorical variables
df3 <- mutate(df, race_des = case_when (race == 1~"a_white", race == 2~"black", race == 3~"other"), 
              smoke_des = case_when(smoke == 0~"no", smoke == 1~"yes"), 
              ht_des = case_when(ht == 0~"no", ht == 1~"yes"), 
              ui_des = case_when(ui == 0~"no",ui == 1~"yes"),
              ptl_comb = case_when (ptl == 0 ~"no", ptl > 0 ~"yes"),
              ftv_comb = case_when (ftv == 0 ~"concern", ftv == 1 ~"a_normal", ftv ==2 ~"a_normal", ftv > 2 ~"concern"))


summary(df3)       
              
## For use with SAS commands since last alphabetical category is used as reference group
df4 <- mutate(df, race_des = case_when (race == 1~"white", race == 2~"black", race == 3~"other"), 
              smoke_des = case_when(smoke == 0~"no", smoke == 1~"yes"), 
              ht_des = case_when(ht == 0~"z_no", ht == 1~"yes"), 
              ui_des = case_when(ui == 0~"z_no",ui == 1~"yes"),
              ptl_comb = case_when (ptl == 0 ~"z_no", ptl > 0 ~"yes"),
              ftv_comb = case_when (ftv == 0 ~"concern", ftv == 1 ~"normal", ftv == 2 ~"normal", ftv > 2 ~"concern"))
```

Biostatistics 213: Homework #5
Due: October 12


Goal: To balance the issues of confounding and collinearity in developing a process for building a multiple regression model.


Assignment:  	
a. Start to build a multiple regression model for infant birthweight, identifying all factors that affect the birthweight of the infant. You should consider all of the covariates as possible predictors. Include in the model all the covariates that are scientifically necessary.



Recall Univariate Results: 

Relationship between mean infant birth weight and mother characteristics

Predictor              | Significant  | Test Used         | p-value
-----------------------|--------------|-------------------|----------
Age                    | No           | Spearman's Corr.  | 0.4037
Weight at last MC      | Yes          | Spearman's Corr.  | 0.0006
Race                   | Yes          | ANOVA             | 0.0068
Smoking Status         | Yes          | T-Test            | 0.0075
Preterm Labor History  | Yes          | T-Test            | 0.0026
Hypertension History   | Yes          | T-Test            | 0.0448
Uterine Irritability   | Yes          | T-Test            | 0.0005
Physician Visits       | Yes          | T-test            | 0.0494


```{r lm with all variables, include = TRUE}

#Assumptions of linear regression: LINE, Linear-relationship among the variables, Independent relationships, N, Equal variance of the residuals

model1 <- bwt ~ age + lwt + race_des + ht_des + smoke_des + ptl_comb + ftv_comb
fit1 <- lm(model1, df3)

summary(fit1)
confint(fit1)

#Akaike’s information criterion
AIC(fit1)

#BIC(Bayesian information criterion
BIC(fit1)

plot(fit1)

#"A large difference between the R-Squared and Adjusted R-squared is not appreciated and generally indicates that multicollinearity exists within the data."
```


```{r lm, model without ftv_comb, include=TRUE}
model2 <- bwt ~ age + lwt + race_des + ht_des + smoke_des + ptl_comb
fit2 <- lm(model2, df3)

summary(fit2)
confint(fit2)

#Akaike’s information criterion
AIC(fit2)

#BIC(Bayesian information criterion
BIC(fit2)


plot(fit2)
```

```{r lm model without ftv_comb and age, include=TRUE}

model3 <- bwt ~ lwt + race_des + ht_des + smoke_des + ptl_comb
fit3 <- lm(model3, df3)

summary(fit3)
confint(fit3)

#Akaike’s information criterion
AIC(fit3)

#BIC(Bayesian information criterion
BIC(fit3)


plot(fit3)
```
```{r lm model without ftv_comb,age, and ptl_comb, include=TRUE}

model4 <- bwt ~ lwt + race_des + ht_des + smoke_des
fit4 <- lm(model4, df3)

summary(fit4)
confint(fit4)

#Akaike’s information criterion
AIC(fit4)

#BIC(Bayesian information criterion
BIC(fit4)

plot(fit4)
```
Model Summaries from manual assmbly
-----------------------------------------------------
model1 - age, lwt, race, smoke, ht, ptl, ftv
RSE       666.6
mR2       0.1995
R2 adj    0.1639
F-stat    5.606
p-value   2.375e-06
AIC       3004.992
BIC       3037.41

model2 - age, lwt, race, smoke, ht, ptl
RSE       665.9
mR2       0.1968
R2 adj    0.1657
F-stat    6.335
p-value   1.152e-06
AIC       3003.621
BIC       3032.797

model3 - lwt, race, smoke, ht, ptl
RSE       664.1
mR2       0.1968
R2 adj    0.1703
F-stat    7.432
p-value   3.953e-07
AIC       3001.621 
BIC       3027.555

model4 - lwt, race, smoke, ht
RSE       670.3
mR2       0.1772
R2 adj    0.1547
F-stat    7.881
p-value   9.606e-07
AIC       3004.181
BIC       3026.873


Of the four models that I manually tested, Model 3 (containing variables lwt, race, smoking status, hypertension, and preterm labor) has the smallest residual standard error and the smallest p-value without losing a favorable R-squared value.



b. Start to build a multiple regression model for infant birthweight, identifying all factors that affect the birthweight of the infant.  You should consider all of the covariates as possible predictors. Use one (or more) of the automated procedures to build an initial model. Consider what steps you would take next.

```{r automated stepwise multiple regression, include=FALSE}

library(olsrr)

autovar1 <- bwt ~ age + lwt + race_des + ht_des + smoke_des + ptl_comb + ftv_comb
automodel1 <- lm(autovar1, df3)

test <- ols_step_both_p(automodel1, pent = 0.1, prem = 0.3, details = TRUE)
plot(test, model = automodel1, print_plot = TRUE)

test$automodel1

```
Results:

Final model summary from Forward Selection
------------------------------------------------------------------
R                       0.444       RMSE                  664.079 
R-Squared               0.197       Coef. Var              22.551 
Adj. R-Squared          0.170       MSE                441000.812 
Pred R-Squared          0.133       MAE                   526.761 
------------------------------------------------------------------
 RMSE: Root Mean Square Error 
 MSE: Mean Square Error 
 MAE: Mean Absolute Error 

                                 ANOVA                                   
------------------------------------------------------------------------
                    Sum of                                              
                   Squares         DF    Mean Square      F        Sig. 
------------------------------------------------------------------------
Regression    19665116.466          6    3277519.411    7.432    0.0000 
Residual      80262147.819        182     441000.812                    
Total         99927264.286        188                                   
------------------------------------------------------------------------


Parameter Estimates                                        
------------------------------------------------------------------------------------------------
        model        Beta    Std. Error    Std. Beta      t        Sig        lower       upper 
------------------------------------------------------------------------------------------------
  (Intercept)    2743.507       248.940                 11.021    0.000    2252.327    3234.688 
  ptl_combyes    -286.668       135.985       -0.144    -2.108    0.036    -554.978     -18.357 
race_desblack    -471.200       149.817       -0.223    -3.145    0.002    -766.802    -175.599 
race_desother    -353.569       115.788       -0.233    -3.054    0.003    -582.028    -125.110 
 smoke_desyes    -339.031       108.252       -0.228    -3.132    0.002    -552.621    -125.441 
          lwt       4.639         1.719        0.195     2.698    0.008       1.246       8.032 
    ht_desyes    -512.705       204.630       -0.172    -2.506    0.013    -916.458    -108.952 
------------------------------------------------------------------------------------------------


```{r Forward and Reverse by AIC, include = FALSE}
test2 <- ols_step_forward_aic(automodel1, pent = 0.1, prem = 0.3, details = TRUE)
plot(test2, model = automodel1, print_plot = TRUE)
summary(test2)

test3 <- ols_step_backward_aic(automodel1, pent = 0.1, prem = 0.3, details = TRUE)
plot(test3, model = automodel1, print_plot = TRUE)
summary(test3)


```

Final model summary from Reverse Selection by AIC
------------------------------------------------------------------
R                       0.444       RMSE                  664.079 
R-Squared               0.197       Coef. Var              22.551 
Adj. R-Squared          0.170       MSE                441000.812 
Pred R-Squared          0.133       MAE                   526.761 
------------------------------------------------------------------
 RMSE: Root Mean Square Error 
 MSE: Mean Square Error 
 MAE: Mean Absolute Error 

                                 ANOVA                                   
------------------------------------------------------------------------
                    Sum of                                              
                   Squares         DF    Mean Square      F        Sig. 
------------------------------------------------------------------------
Regression    19665116.466          6    3277519.411    7.432    0.0000 
Residual      80262147.819        182     441000.812                    
Total         99927264.286        188                                   
------------------------------------------------------------------------

                                      Parameter Estimates                                        
------------------------------------------------------------------------------------------------
        model        Beta    Std. Error    Std. Beta      t        Sig        lower       upper 
------------------------------------------------------------------------------------------------
  (Intercept)    2743.507       248.940                 11.021    0.000    2252.327    3234.688 
          lwt       4.639         1.719        0.195     2.698    0.008       1.246       8.032 
race_desblack    -471.200       149.817       -0.223    -3.145    0.002    -766.802    -175.599 
race_desother    -353.569       115.788       -0.233    -3.054    0.003    -582.028    -125.110 
    ht_desyes    -512.705       204.630       -0.172    -2.506    0.013    -916.458    -108.952 
 smoke_desyes    -339.031       108.252       -0.228    -3.132    0.002    -552.621    -125.441 
  ptl_combyes    -286.668       135.985       -0.144    -2.108    0.036    -554.978     -18.357 
