---
title: "BST213 Homework #2"
author: "Rebecca Kelly"
date: "2022-09-07"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(fig.width = 4, fig.asp = 0.618, echo = FALSE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(tableone)
library(confintr)
library(haven)
```

```{r data import, include=FALSE}

#Import dataset
df<-(read_csv("lbw.csv"))
summary(df)
View(df)
df2 <- mutate(df, race_des = case_when(race == 1~"white", race == 2~"black", 
                                       race == 3~"other"), 
              smoke_des = case_when(smoke == 0~"No", smoke == 1~"Yes"), 
              ht_des = case_when(ht == 0~"No", ht == 1~"Yes"), 
              ui_des = case_when(ui == 0~"No",ui == 1~"Yes"))
              options(repr.plot.width =2, repr.plot.height =2)
```

Goals:    To be able to interpret simple linear regression models with 
          continuous, binary predictor, or categorical predictors.


Assignment: 


Simple Linear Regression Results

Predictor  
------------------------------------------------------------
continuous    - *Weight at last MC*     (#1)         
binary        - *Hypertension History*  (#2)          
categorical   - *Race*                  (#3)              


Assumptions for Simple Linear Regression: 

Linear Relationship between the predictor and the outcome
All data points are independent
Normal distribution of the residuals
Equal variance of the residuals 



#1. Run a model that includes maternal weight as a continuous predictor 
of birth weight. Interpret the regression portion (i.e., the bottom) of 
the printout. Compare the regression results to a Pearson correlation 
coefficient. Use PROC GLM to run the same model. Interpret and compare 
the results to the regression results from above.


 
```{r, lab9 code bwt v lwt simple linear regression, include = TRUE}

df2 %>%
  ggplot(aes(x=lwt,y=bwt)) +
  geom_point(alpha=0.7) +
  labs(x="Weight of Mother (lbs)", y="Birthweight (grams)")

mod2 <- lm(lwt ~ bwt, df2)
summary(mod2)
pred2 <- predict(mod2, interval="prediction")
newdf2 <- cbind(df2, pred2)
```

```{r, bwt v lwt linear reg, include = TRUE}
df2 %>%
  ggplot(aes(x=lwt,y=bwt)) +
  geom_point(alpha=0.95) +
  labs(x="Weight of Mother (lbs)", y="Birthweight (grams)")+
  geom_smooth(method = "lm", se = TRUE)
```

```{r, plot, include = FALSE}
plot(mod2, 1)
```

```{r histogram, include = TRUE}

ggplot(data = df2, aes(x = mod2$residuals)) +
geom_histogram(bins = 25) +
labs(title = 'Histogram of Residuals', x = 'Residuals', y = 'Frequency')

```


```{r calculate the CI for beta, include=FALSE}
#Calculate confidence interval
predict(mod2, interval="confidence")

```

Results from simple linear regression: 

y = a + bx
y = 106.8 + 0.007820x
p-value = 0.01021
R2 = 0.02959
 
Interpretation: 
The average weight of the infant changes by 0.0078290 grams for each one 
pound change in weight of the mother. The infant weighs 106.8 grams if the
mother weighs 0 pounds. The variable weight of the mother accounts for 
0.02959 of the total variability in infant weight. 



Weight of mother at last menstrual cycle

Type: Continuous
Determine Normality: Histogram



Test of mother's weight at last menstrual cycle with birth weight:
 
H0: There is no correlation between mother's weight at last menstrual cycle 
and infant birth weight. (r = 0) 
  
HA: There is a correlation between mother's weight at last menstrual cycle 
and infant birth weight. (r does not equal 0)

```{r compute pearson coefficient lwt bwt, fig.align = "center", include = TRUE}

cor.test(df2$lwt, df2$bwt, method= "pearson")

ci_cor(df2$lwt, df2$bwt,
  probs = c(0.025, 0.975),
  method = c("pearson"))

```

Results from simple linear regression: 

y = a + bx
y = 106.8 + 0.007820x
p-value = 0.01021

Results from pearson correlation: 

0.1864291 (0.044893, 0.320627)
p-value = 0.01021


Compare results: 
Both methods show a positive, linear relationship between infant birth
weight and mother's weight. 
The p-value from each approach is the same: 0.01021. 


***

#2.Run a model that includes history of hypertension as a binary predictor 
of birth weight. Interpret the regression portion of the printout. Compare 
the regression results to a 2-sample t-test. Use PROC GLM to run the same 
model. Interpret and compare the results to the regression results from 
above.


```{r, lab9 code ht v bwt simple linear regression, include=TRUE}

df2 %>%
  ggplot(aes(x=ht ,y=bwt)) +
  geom_boxplot(aes(x = factor(ht), y = bwt,
      fill = factor(ht))) +
  labs(x= "Birthweight (grams)", y="History of Hypertension")

mod3 <- lm(bwt ~ ht, df2)
summary(mod3)


pred3 <- predict(mod3, interval="prediction")
newdf3 <- cbind(df2, pred3)
```

```{r, ht v bwt linear reg, echo=FALSE}
newdf3 %>%
  ggplot(aes(ht, bwt))+
  labs(x="History of Hypertension",y= "Birthweight (grams)") +
  geom_point()+
  geom_smooth(method = "lm", se = TRUE)
  
```

```{r, include=TRUE}
ggplot(data = df2, aes(x = mod3$residuals)) +
  geom_histogram(bins = 25) +
  labs(title = 'Histogram of Residuals', x = 'Residuals', y = 'Frequency')
```

Linear Regression Results: 

y = a + bx
y = 2972 + -436 x
p-value = 0.0448
R2 value = 0.01611
 
Interpretation: For mothers without hypertension, the mean birth weight of
infants is 2972 grams.For mothers with hypertension, the mean birth weight of
infants decreases by 436 grams. Mother's hypertension explains 0.016 of 
the variation in infant birth weight.  


T-Test:
```{r independent 2-sample t-test ht bwt, fig.align = "center", include=TRUE}

t.test(bwt ~ ht, data = df2, alternative = c("two.sided"), mu = 0, 
       paired = FALSE, var.equal = TRUE)
```
```{r difference in means, include=FALSE}
2972.424 - 2536.750

```
2972.424 - 2536.750 = 435.674

435.674 (10.117, 861.23)

Conclusion from univariate test: Based on the p-value of 0.04485, I 
reject the null hypothesis that there is no difference in mean birth weights 
of infants born to mothers who had hypertension compared to mothers who did 
not have hypertension. There is  a statistically significant difference 
in mean birth weight between mothers with hypertension and those without 
hypertension. 


Comparison of methods: 

The p-value and the differences are the same for both approaches. 



***

3. Use PROC REG to run a model that includes race (in 3 categories) as 
a categorical predictor of birth weight. Interpret the regression portion 
of the printout. Compare the regression results to a standard ANOVA analysis.  
Use PROC GLM to run the same model.  Interpret and compare the results to 
the regression results from above.

H0: There is no relationship between race and birthweight (x1 = 0, x2 = 0)
HA: At least one of the race categories is different (X1 or X2 !=0)

```{r, lab9 code bwt v race, simple linear regression, include = TRUE}


mod4 <- lm(bwt ~ race_des, df2)
summary(mod4)

df2 %>%
  ggplot(aes(x=race_des,y=bwt)) +
  geom_point(alpha=0.95) +
  labs(x="Race", y="Birthweight (grams)") +
  geom_smooth(method = "lm", se = TRUE)
```

```{r residuals, include = FALSE}
pred4 <- predict(mod4, interval="prediction")
newdf4 <- cbind(df2, pred4)

plot(mod4,1)
```

```{r}
ggplot(data = df2, aes(x = mod4$residuals)) +
  geom_histogram(bins = 25) +
  labs(title = 'Histogram of Residuals', x = 'Residuals', y = 'Frequency')
```





This model uses race = black as the reference group due to the 
alphabetical arrangement of my race_des column by categories. 

Results from simple linear regression: 

y = a + bx
y = 2719.69 + x1(84.32) + x2(384.26) 
p-value = 0.6099, 0.0159
R2 = 0.0406

Interpretation: Infants born to black mothers have an average birth weight
of 2719.69 grams. Infants born to mothers of other races have an average
birth weight of 2804.01 grams. Infants born to white mothers have an average
birth weight of 3103.95 grams. Based on the p-value of 0.0159 for the
x2 coefficient, infants born to white mothers have a higher mean birth
weight from black or other mothers. 
other mothers. 


Approach: 
H0: The mean birth weight of infants is the same among different racial groups.

HA: The mean birth weight of infants in at least one of the race categories is 
different from the others. 

Test of significance: Analysis of Variance 
```{r Analysis of Variance bwt race, fig.align = "center", include=TRUE}
anova <- aov(bwt ~ race_des, df2)
summary(anova)

#Multiple testing

pairwise.t.test(df2$bwt, df2$race_des, p.adjust.method="bonferroni")

```

Comparison of methods: 

Both methods conclude that there is a positive relationship between
average birthweight and infants born to white mothers versus black and other
mothers. 
 