```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 4, fig.asp = 0.618, echo = FALSE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(tableone)
library(confintr)
```

```{r data import, include=FALSE}
#Import dataset
df<-(read_csv("lbw.csv"))
summary(df)
View(df)
df2 <- mutate(df, race_des = case_when(race == 1~"white", race == 2~"black", 
                                       race == 3~"other"), 
              smoke_des = case_when(smoke == 0~"No", smoke == 1~"Yes"), 
              ht_des = case_when(ht == 0~"No", ht == 1~"Yes"), 
              ui_des = case_when(ui == 0~"No",ui == 1~"Yes"))
              options(repr.plot.width =2, repr.plot.height =2)
```
 
Goals:	Remind you of the simple univariate tests you have learned previously;
        Set the groundwork for building multiple regression models;
        Practice with correlation coefficients and scatter plots.

Assignment:  Using the data set on birth weights, use the continuous measure 
             of birth weight, BWT, as the outcome (not the dichotomized one) 
             and:
             

*1. Find all the univariate predictors of birth weight.  Do not use regression. 
Use “simple” tests, like t-tests, Wilcoxon tests, ANOVA, correlations, etc.*


### Summary of Results


## Relationship between mean infant birth weight and mother characteristics


Predictor              | Significant  | Test Used         | p-value
-----------------------|--------------|-------------------|----------
Age                    | No           | Spearman's Corr.  | 0.4037
Weight at last MC      | Yes          | Spearman's Corr.  | 0.0006
Race                   | Yes          | ANOVA             | 0.0068
Smoking Status         | Yes          | T-Test            | 0.0075
Preterm Labor History  | Yes          | Kruskal Wallis    | 0.0093
Hypertension History   | No           | T-Test            | 0.1329
Uterine Irritability   | Yes          | T-Test            | 0.0005
Physician Visits       | No           | ANOVA             | 0.4230

Note:
I am disregarding the variables "Identification Code" and "Low Birth 
Weight(diochotomous outcome)" due to biological irrelevance and redundancy, respectively.


## Outcome - Dependent Variable : Birth Weight (Continuous)

Type: Continuous
Determine Normality: Histogram
      
```{r bwt histogram, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_histogram(aes(x = bwt), binwidth = 500 )+
      labs(title = "Distribution of Birth Weight", x = "Birth Weight (grams)") +
      theme_classic()
```


Conclusion: The distribution of Birth Weights is approximately normal and 
does not require any transformation prior to subsequent analysis. 

***

## Covariate 1 : Mother's Age

Type: Continuous
Determine Normality: Histogram
      
```{r age histogram, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_histogram(aes(x = age), binwidth = 5 ) +
      labs(title = "Distribution of Mothers Age", x = "Age (years)") +
      theme_classic()
```


Conclusion: The mother's age distribution is skewed to the right. We could 
either transform the data or use non-parametric methods to analyze the data.
I will use a non-parametric approach to keep things simple.
    
Inspect scatter plot: 
```{r scatter plot of age vs bwt, fig.align = "center", echo=FALSE}

df2 %>% ggplot() +
  geom_point(aes(x = age, y = bwt)) +
  labs(title = "Mother's Age v.s. Infant Birthweight",
       x = "Mother's Age (years)", y = "Birthweight (grams)") +
  theme_bw()
```

  
Test of mother's age relationship with birth weight:
H0 = There is no correlation between mother's age and infant birth weight. 
(r = 0) 
HA = There is a correlation between mother's age and infant birth weight. 
(r does not equal 0)
```{r compute spearman coefficient age bwt, fig.align = "center", echo=FALSE}

cor.test(df2$age, df2$bwt, method= "spearman")
```
Conclusion for covariate 1: I fail to reject the null hypothesis that mother's 
age and infant birth weight are not correlated. Based on a p-value of 0.4037, 
there is no linear relationship between mother's age and infant birth weight. 
  
***

## Covariate 2 : Weight of mother at last menstrual cycle

Type: Continuous
Determine Normality: Histogram
```{r histogram lwt, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_histogram(aes(x = lwt), binwidth = 20 ) +
      labs(title = "Distribution of Mother's Weight at Last Menstrual Cycle", x = "Weight (lbs)")       +
      theme_classic()
```


Approaches:The mother's weight at last menstrual cycle age distribution is 
skewed to the right. We could either transform the data or use non-parametric 
methods to analyze the data. I will use a non-parametric approach to keep 
things simple. 
    
Inspect Scatter Plot: 
```{r scatter plot of lwt vs bwt, fig.align = "center", echo=FALSE}

      df2 %>% ggplot() +
      geom_point(aes(x =lwt, y =bwt)) +
      labs(title = "Mother's Weight at last Menstrual Cycle v.s. Infant Birthweight",
      x = "Weight (lbs)", y = "Birthweight (grams)") +
      theme_bw()
```
Test of mother's weight at last menstrual cycle with birth weight:
 
H0: There is no correlation between mother's weight at last menstrual cycle 
and infant birth weight. (r = 0) 
  
HA: There is a correlation between mother's weight at last menstrual cycle 
and infant birth weight. (r does not equal 0)
```{r compute spearman coefficient lwt bwt, fig.align = "center", echo=FALSE}

cor.test(df2$lwt, df2$bwt, method= "spearman")

ci_cor(df2$lwt, df2$bwt,
  probs = c(0.025, 0.975),
  method = c("spearman"),
  type = c("bootstrap"),
  boot_type = c("bca"),
  R = 999,
  seed = 1)
```


Conclusion for Covariate 2: I reject the null hypothesis that mother's weight 
at last menstrual period and infant birth weight are not correlated. Based on 
a p-value of 0.00056, there is a linear relationship between mother's weight 
at last menstrual period and infant birth weight. 

***

## Covariate 3 : Race

Type: Categorical; 1 = white, 2 = black, 3 = other
Determine Normality: Boxplots
```{r boxplot bwt race, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_boxplot(aes(x = factor(race_des), y = bwt,
      fill = factor(race_des))) +
      labs(title = "Birthweight by Race", x = "Race", y = "Birthweight (grams)") +
      theme_minimal()
```


Approach: 
H0: The mean birth weight of infants is the same among different racial groups.
HA: The mean birth weight of infants in at least one of the race categories is 
different from the others. 

Test of significance: Analysis of Variance 
```{r Analysis of Variance bwt race, fig.align = "center", echo=FALSE}
anova <- aov(bwt ~ race, df2)
summary(anova)
```
Based on an F-statistic of 7.50 corresponding to a p-value of 0.00678, I reject
the null hypothesis that there is no difference in mean birth weight among 
infants in different racial groups. 

To identify which groups may be different from one another, I can perform 
pair-wise t-tests. 
```{r pairwise t-test bwt race, fig.align = "center", echo=FALSE}
pairwise.t.test(df2$bwt, df2$race_des, p.adjust.method="bonferroni")
```
Conclusions: Based on pair-wise t-tests of the three racial categories 
compared, there are differences in mean birth weight between infants born 
to white and black mothers, and white and other mothers. 

***

## Covariate 4 : Smoking Status during Pregnancy

Type: Dichotomous
Determine Normality: Boxplots

```{r boxplot smoke stat, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_boxplot(aes(x = factor(smoke_des), y = bwt,
      fill = factor(smoke_des))) +
      labs(title = "Birthweight by Smoking Status", x = "Smoking", y = "Birthweight (grams)") +
      theme_minimal()
```


Approach: The distributions are relatively normal, so I will use parametric 
methods. I will perform an independent t-test between the two groups to 
determine whether or not the means of the groups are the same. 

H0: There is no difference in mean birth weight of infants born to mothers who 
smoked compared to mothers who did not smoke. (meu1 - meu2 = 0)
HA: There is a difference in mean birth weight of infants born to mothers who 
smoked compared to mothers who did not smoke. (meu1 - meu2 != 0)


Test of significance:Unpaired t-test
```{r independent 2-sample t-test bwt smoke, fig.align = "center", echo=FALSE}

t.test(df2$bwt~df2$smoke_des)# where y is numeric and x is a binary factor

```
Conclusion: Based on the t-statistic of 2.71, corresponding to a p-value 
of 0.0075, I reject the null hypothesis that there is no difference between 
the mean birth weight of infants born to mothers who are non-smokers and to 
mothers who are smokers. 
      
***

## Covariate 5 : History of Preterm Labor

Type: Dichotomous
Determine Normality: Boxplots
```{r boxplot ptl bwt, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_boxplot(aes(x = factor(ptl), y = bwt,
      fill = factor(ptl))) +
      labs(title = "Birthweight by History of Premature Labor", x = "Number of Premature Labors", y = "Birthweight (grams)") +
      theme_minimal()
```


Approach: Three of the distributions appear skewed. I will need to use 
non-parametric methods to determine if one of the groups is different 
from the others. 

Test of significance: Kruskal Wallis 

H0: The median of the distributions are all at the same location.
HA: The median of at least one distribution is at a different location. 
```{r kruskal wasllis test bwt ptl, fig.align = "center", echo=FALSE}
kruskal.test(bwt ~ ptl, data = df2)

pairwise.wilcox.test(df2$bwt, df2$ptl,
                 p.adjust.method = "BH")

```

Conclusion: Based on the p-value from the Kruskal Wallis test (0.0093), I 
reject the null hypothesis that the median birth weight was the same among 
women with different numbers of previous preterm labors. The multiple 
comparisons Wilcoxon rank sum test suggests that it is birth weight of 
infants to mothers with one prior pre-term birth is different from the others. 

***

## Covariate 6 : History of Hypertension

Type: Dichotomous
Determine Normality: 
```{r boxplot ht bwt, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_boxplot(aes(x = factor(ht_des), y = bwt,
      fill = factor(ht_des))) +
      labs(title = "Birthweight by History of Hypertension", x = "History of Hypertension", y = "Birthweight (grams)") +
      theme_minimal()
```


Approach: The distributions appear normal. I can use parametric methods 
for a test of significance. 

Test of significance: independent, 2-sample t-test

H0: There is no difference in mean birth weights of infants born to mothers 
who had hypertension compared to mothers who did not have hypertension. 
(meu1 - meu2 = 0)
HA: There is a difference in mean birth weights of infants born to mothers 
who had hypertension compared to mothers who did not have hypertension. 
(meu1 - meu2 != 0)

```{r independent 2-sample t-test ht bwt, fig.align = "center", echo=FALSE}

t.test(df2$bwt~df2$ht_des)# where y is numeric and x is a binary factor

```
Conclusion: Based on the p-value of 0.133, I fail to reject the null 
hypothesis that there is no difference in mean birth weights of infants born 
to mothers who had hypertension compared to mothers who did not have 
hypertension. There is not a statistically significant difference in mean 
birth weight between mothers with hypertension and those without hypertension. 

***

## Covariate 7 : Presence of Uterine Irritability 

Type: Dichotomous
Determine Normality: 
```{r boxplot ui bwt, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_boxplot(aes(x = factor(ui_des), y = bwt,
      fill = factor(ui_des))) +
      labs(title = "Birthweight by Uterine Irritability", x = "History of Uterine Irritability", y = "Birthweight (grams)") +
      theme_minimal()
```


Approach: The distributions appear normal. I can use parametric methods for 
a test of significance. 

Test of significance: independent 2-sample t-test

H0: There is no difference in mean birth weight of infants born to mothers 
who had uterine irritability compared to mothers who did not have uterine 
irritability. (meu1 - meu2 = 0)
HA: There is a difference in mean birth weight of infants born to mothers 
who had uterine irritability compared to mothers who did not have uterine 
irritability. (meu1 - meu2 != 0)

Test of significance:
```{r independent 2-sample t-test bwt ui, fig.align = "center", echo=FALSE}

t.test(df2$bwt~df2$ui_des)# where y is numeric and x is a binary factor

```
Conclusion: Based on the p-value of 0.00047, I reject the null hypothesis 
that there is no difference between mean birth weight of infants born to 
mothers with uterine irritability compared to mothers without uterine 
irritability. There is statistical support for the hypothesis that 
mother's uterine irritability is related to infant birth weight. 

***

## Covariate 8 : Number of Physician Visits during the First Trimester 

Type: Count
Determine Normality: 
```{r boxplot bwt ftv, fig.align = "center", echo=FALSE}
      df2 %>% ggplot() +
      geom_boxplot(aes(x = factor(ftv), y = bwt,
      fill = factor(ftv))) +
      labs(title = "Number of Physician Visits in 1st Trimester versus Infant Birth Weight", x = "Number of Physician Visits", y = "Birthweight (grams)") +
      theme_minimal()
```


Approach: The distribution of birthweight among the number of physician visit     categories, so I will proceed with parametric tests. 

Test of significance: Analysis of Variance (ANOVA) 
```{r Analysis of Variance bwt ftv, fig.align = "center", echo=FALSE}
anova <- aov(bwt ~ ftv, df2)
summary(anova)
```
Conclusions: Based on an F-statistic of 0.644 corresponding to a p-value 
of 0.423, I fail to reject the null hypothesis that there is no difference 
in mean birth weight among infants whose mothers had a different number of 
physician visits during the first trimester. 

***




*2. Use scatter plots and correlation coefficients to look at the 
relationship between mother's age versus birth weight and, separately, 
mother’s weight versus birth weight. (Note: You may have done these 
already for #1 above.)*



See results from #1.



